---
title: "Activity: Street or Road" 
author: "Eli Freedman"
date: "11/16/2021"
output: html_notebook
---

## Front Matter

```{r}
# clean up work space environment
rm(list = ls())

# all packages used for the assignment
library(tidyverse)
library(mosaic)
library(dcData)
library(DataComputing)

# load all data sets
download.file(url="https://mdbeckman.github.io/dcSupplement/data/CMS_ProvidersSimple.rds",
              destfile = "CMS_ProvidersSimple.rds")
CMS_ProvidersSimple <- readRDS("CMS_ProvidersSimple.rds")
StreetAddresses <- 
  readr::read_csv("https://mdbeckman.github.io/dcSupplement/data/street-addresses.csv")
```

## Pre-Activity Questions

How many addresses can be found in both data sets?
```{r}
full_join(CMS_ProvidersSimple, StreetAddresses, by = NULL) %>%
  summarize(numAddresses = n())
```

## Problem 18.14

Pull a sample from the larger `StreetAddresses` data set
```{r}
StreetAddresses_Sample <-
  StreetAddresses %>%
  sample_n(size = 50)

StreetAddresses_Sample
```

In everyday language, describe a pattern that you think will identify the information you are looking for.
1. The PO Box cases tend to have a sub string “PO”.

Translate (1) into the form of a regular expression.
2. The regular expression for “PO” is simply "PO".

Filter to retain the cases that match the expression. Hint: filter() and grepl() are useful for this.
3. Find some cases that match:
```{r}
Matches <- 
  StreetAddresses_Sample %>%
  filter(grepl(pattern = "PO", address))

Matches
```

Filter to retain the cases that do not match the expression.
4. Find cases that don’t match:
```{r}
Dont <- 
  StreetAddresses_Sample %>%
  filter(!grepl(pattern = "PO", address))

Dont
```

Examine the results of (3) and (4) to identify shortcomings in your patterns.
5. Find any cases in the `Matches` that shouldn’t be there (none so far in the excerpt shown). Find any cases in `Dont` that should have matched, for example we want to try and capture the “NCSU BOX” results among our matches.

Improve or extend the pattern to deal with the mistaken cases.
6. It looks like “BOX” is a better pattern. Since the box number is wanted, the regex should include an identifier for the number inside extraction parentheses.
```{r}
pattern <- "BOX\\s+(\\d+)"

Matches <- 
  StreetAddresses_Sample %>% 
  filter(grepl(pattern, address))

Dont <-
  StreetAddresses_Sample %>%
  filter(!grepl(pattern, address))

Dont
```

Repeat until satisfied.
7. The result seems satisfactory

Put extraction parentheses around the parts of the regular expression that contain the info you want.
8. Use tidyr::extract() to pull out the part of the pattern identified by extraction parentheses.
```{r}
BoxNumbers <- 
  StreetAddresses_Sample %>%
  filter(grepl(pattern, address)) %>%
  tidyr::extract(address, into = "boxnum", regex = pattern)

BoxNumbers
```

## Problem 18.15

1. Explain each line of code
```{r}
pattern <- "(ST|RD|ROAD)"                            #Create a new variable `pattern` which is equal to one of the three of the following strings: 'ST', 'RD', or 'ROAD'
LeftOvers <-                                         #Create a new data frame `LeftOvers` which is equal to the remaining commands
  StreetAddresses %>%                                #Loads the `Addresses` data set (I named this `StreetAddresses`)
  filter( !grepl(pattern, address),                  #Filters for cases that do not contain the pattern
          !grepl("\\sAPT|UNIT\\s[\\d]+$", address),  #Filters for cases that do not contain 'APT' or 'UNIT'
          !grepl(" BOX ", address)                   #Filters for cases that do not contain 'BOX'
          )                                          #End of Filter
LeftOvers                                            #Call the data frame `LeftOvers`
```

2. Expand to match at least 12 patterns
```{r}
pattern_new <- "(ST|RD|ROAD|LN|LANE|CIR|WAY|TRAIL|DR|DRIVE|AVE|CT|COURT|BLVD|POINT|PL|PLACE|PATH)"
LeftOvers_new <-
  StreetAddresses %>%
  filter(!grepl(pattern_new, address),
         !grepl("\\sAPT|UNIT\\s[\\d]+$", address),
         !grepl(" BOX ", address)
         )
LeftOvers_new
```

3. Provide a table in descending order of popularity for the street name identifiers
```{r}
popularIdentifiers <-
  full_join(CMS_ProvidersSimple, StreetAddresses, by = 'address') %>%
  summarize(address) %>%
  filter(grepl(pattern_new, address)) %>%
  tidyr::extract(address, into = "identifier", regex = pattern_new) %>%
  group_by(identifier) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

popularIdentifiers
```

4. Construct a bar chart for the previous exercise
```{r}
options(scipen = 100) #Remove scientific notation

ggplot(data = popularIdentifiers, aes(x = reorder(identifier, -count), y = count)) +
  geom_bar(stat = 'identity')
```

5. It appears that in the joined data frame of `CMS_ProvidersSimple` and `StreetAddresses`, the most popular street ending (by far) is 'ST'. The least common street ending is 'PATH'.