---
title: "HELPrct (Data Verbs) Activity"
author: "Eli Freedman"
date: "9/28/2021"
output: html_notebook
---


## Set Up:

```{r}
# load packages
library(tidyverse)  # for data verbs like `group-by()` & `summarise()`
library(mosaicData) # includes `HELPrct` data set 
library(mosaic)     # includes `mplot()` interactive plot builder
library(esquisse)   # includes `esquisser()` interactive plot builder


# Load the HELPrct data from `mosaicData` into our RStudio environment
data("HELPrct", package = "mosaicData")

```


## Data Description

*TASK 1: Write a couple of sentences describing the `HELPrct` data--What is the setting for the data? What are they about?* 

The 'HELPrct' data set is a clinical trial for adult inpatients taken from a detoxification unit. This means that each case from the data set describes the information of a patient that was under some sort of intoxication.

## Basic Summaries with `summarise()` 

*TASK 2: Produce an expression involving `summarize()` and `HELPrct` that will produce the following.*

1. *number of people (cases) in `HELPrct` study*
2. *combined total number of times in the past 6 months all the people in `HELPrct` entered a detox program (measured at baseline)* 
3. *mean time (in days) to first use of any substance post-detox for all the people in `HELPrct`*


```{r}
Task2 <-
  HELPrct %>%
  filter(e2b != 'NA', daysanysub != 'NA') %>%
  summarise(cases = n(),
            totalDetox = sum(e2b),
            meanFirstSubstance = mean(daysanysub))

Task2
```


## Subgrouping with `group_by()` 

*TASK 3: Repeat task 2 above, but calculate the results group-by-group.  Show your code and write a sentence or two about what you observe in the results for each of the following:* 

1. *males versus females*

From the table below, it appears that there are significantly more males than females that are in this case study of patients from a detoxification unit. We can also see that men are more likely to use a substance again sooner than women are.

```{r}
Task3Part1 <-
  HELPrct %>%
  filter(e2b != 'NA', daysanysub != 'NA') %>%
  group_by(sex) %>%
  summarise(cases = n(),
            totalDetox = sum(e2b),
            meanFirstSubstance = mean(daysanysub))
Task3Part1
```


2. *homeless or not*

From the table below, we can see that in this sample case study, there are more homeless substance abusers in the detoxification unit than those that have a place of residence.

```{r}
Task3Part2 <-
  HELPrct %>%
  filter(e2b != 'NA', daysanysub != 'NA') %>%
  group_by(homeless) %>%
  summarise(cases = n(),
            totalDetox = sum(e2b),
            meanFirstSubstance = mean(daysanysub))
Task3Part2
```

3. *substance*

From the table below, we can see that in this case study, alcohol is the leading substance causing the patient to be in the detoxification unit.

```{r}
Task3Part3 <-
  HELPrct %>%
  filter(e2b != 'NA', daysanysub != 'NA') %>%
  group_by(substance) %>%
  summarise(cases = n(),
            totalDetox = sum(e2b),
            meanFirstSubstance = mean(daysanysub))
Task3Part3
```

4. *break down the homeless versus housed further, by sex*

From the table below, we can see that there are more males than females, both housed and homeless, that are in this case study's detoxification unit.

```{r}
Task3Part4 <-
  HELPrct %>%
  filter(e2b != 'NA', daysanysub != 'NA') %>%
  group_by(homeless, sex) %>%
  summarise(cases = n(),
            totalDetox = sum(e2b),
            meanFirstSubstance = mean(daysanysub))
Task3Part4
```

5. *break down the homeless versus housed further, by substance*

From the table below, we can see that alcohol is the leading cause of substance to bring someone into the detoxification unit in this case study. However, the leading cause for those that are housed is heroin.

```{r}
Task3Part5 <-
  HELPrct %>%
  filter(e2b != 'NA', daysanysub != 'NA') %>%
  group_by(homeless, substance) %>%
  summarise(cases = n(),
            totalDetox = sum(e2b),
            meanFirstSubstance = mean(daysanysub))
Task3Part5
```


## TASK 4:

*Include one or more interesting plots of the data involving at least 3 variables per plot. Write a few sentences to explain the story that your plot tells about these data.  You can use one of the relationships that you studied in Task 2, or you can explore a different group of variables in the `HELPrct` that show something interesting.*  

This plot shows the relationship between age and number of hospitalizations, while the color represents the number of days before they used that substance again. This plot shows that age is not a very contributing factor in relation to number of hospitalizations. It also shows that those with more hospitalizations are more likely to use the substance again in a shorter amount of time comparatively.

```{r}
gf_point(hospitalizations ~ age, data = HELPrct, color = ~ daysanysub) %>% 
  gf_theme(legend.position = "right") %>% 
  gf_labs(title = "", caption = "")
```


