---
title: "Activity: Statistics of Gene Expression"
author: "Eli Freedman"
date: "11/9/2021"
output: html_notebook
---

# Set Up

### Clean up environment and load packages

Here, I call all the libraries I know I will need throughout this activity.

```{r}
# clean up the RStudio environment
rm(list = ls())

# load all packages here
library(tidyverse)
library(mosaic)
library(dcData)
library(DataComputing)
```

# 18.1

### Set Up Data Frames

Here, I set up the data frames. I also call the `SummaryStats` and utilize ggplot to create a bar graph.

```{r}
Narrow <-
  NCI60 %>%
  tidyr::pivot_longer(cols = -Probe, names_to = "cellLine", values_to = "expression")

CellTypes <-
  NCI60cells %>%
  select(cellLine, tissue) %>%
  mutate(cellLine = gsub(pattern = "\\:", replacement = ".", x = as.character(cellLine)))

Narrow <- 
  Narrow %>%
  inner_join(CellTypes)

Probe_TOP3A <-
  Narrow %>%
  filter(Probe == "TOP3A")

SummaryStats <-
  Probe_TOP3A %>%
  group_by(tissue) %>%
  summarise(mn_expr = exp(mean(expression, na.rm = TRUE)))

SummaryStats %>% 
  ggplot(aes(x = tissue, y = mn_expr)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

# 18.2

Here, I call `Probe_TOP3A` to create a scatter plot.

```{r}
Probe_TOP3A %>% 
ggplot(aes(x = tissue, y = exp(expression))) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

# 18.3

Here, I alter `SummaryStats` and then create an overlapping bar graph and scatter plot.

```{r}
SummaryStats <-
  Probe_TOP3A %>%
  group_by(tissue) %>%
  summarise(mn_expr = mean(expression,  na.rm = TRUE),
            se = sd(expression, na.rm = TRUE) / sqrt(n())) 

SummaryStats %>% 
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_bar(stat = "identity", fill = "gray", color = NA) +
  geom_point(data = Probe_TOP3A, aes(x = tissue, y = exp(expression))) + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

# 18.4

Here, I alter `SummaryStats` again, and then create an overlapping bar graph and error bar.

```{r}
SummaryStats <- 
  SummaryStats %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 
SummaryStats %>%
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_bar(stat = "identity", alpha = 0.2) +
  geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

# 18.5

Here, I call `SummaryStats` and `Probe_TOP3A` to create an overlapping scatter plot and error bar.

```{r}
SummaryStats %>%
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_point(data = Probe_TOP3A, aes(x = tissue, y = exp(expression))) +
  geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5)
```

# 18.6

Here, I create the `r2`, `ProbeR2`, and `Actual` data tables. I then use `Actual` to create a scatter plot.

```{r}
r2 <- function(data) {
  mosaic::rsquared(lm(data$expression ~ data$tissue))
}

ProbeR2 <- 
  Narrow %>% 
  group_by(Probe) %>%
  dplyr::do(probe_rsq = r2(.)) %>%
  mutate(r2 = unlist(probe_rsq))

Actual <-
  ProbeR2 %>%
  arrange(desc(r2)) %>%
  head(30) 

Actual %>%
  ggplot(aes(x = reorder(Probe, desc(r2)), y = r2)) +
  geom_point() + 
  xlab("Probe") + 
  ylab("R-squared") + 
  ggtitle("Probes with large R-squared for expression level explained by tissue type") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Here, I specify the Probe with the highest R^2 value: A_23_P61317. After plotting it, I can see that it has a lower expression value compared to the graph in 18.5.

```{r}
Probe_A_23_P61317 <-
  Narrow %>%
  filter(Probe == 'A_23_P61317')

SummaryStats %>%
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_point(data = Probe_A_23_P61317, aes(x = tissue, y = exp(expression))) +
  geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5)
```

# 18.7

Here, I create a new data frame, `NullR2`, and create two overlapping density graphics, one of `ProbeR2` and one of `NullR2`.

```{r}
NullR2 <-
  Narrow %>%
  group_by(Probe) %>%
  mutate(expression = mosaic::shuffle(expression)) %>%
  group_by(Probe) %>%
  do(r2 = r2(.)) %>%
  mutate(r2 = unlist(r2))

ProbeR2 %>%
  ggplot(aes(x = r2)) + 
  geom_density(fill = "gray30", color = NA) +
  geom_density(data = NullR2, aes(x = r2), 
               fill = "gray80", alpha = .75, color = NA)
```

# 18.8

Here, I create the `Null` data frame, and then append it to the `Actual` data frame. I then create two overlapping scatter plots which describe the R^2 values of the Probes in the new `Actual` data frame.

```{r}
Null <-
  NullR2 %>% 
  arrange(desc(r2)) %>%
  head(30) 

# append the 30 highest `Null` values to the `Actual` data
Actual$null <- Null$r2

Actual %>%
  ggplot(aes(x = reorder(Probe, desc(r2)), y = r2)) +
  geom_point() + 
  geom_point(aes(y = null), color = "gray50") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```